### 2018-12-01 report/report_draft.Rmd
added rough outline for the report

---
title: "Identifying Drawings with the Quick Draw Challenge"
author: "Blue Team"
output: 
  pdf_document:
    keep_tex: yes
bibliography: references.bib
---

We're trying to put together a CNN based on the "Quick, Draw!" Kaggle
competition [@QuickDra10].

# Theory & Literature Review

CNN with dilation is preferable to LSTM because it's presumably
faster. We can cite some papers here.

# Hypotheses

Maybe we should structure our models as experiments. Then talk 
about what worked the best. Here's some example hypotheses

$H_a:$ Prediction will go up when we use a CNN compared to a fully
	connected feedforward neural network.

$H_a:$ Prediction will go up when we use regularization

# Data

We got our data from Google/Kaggle

# Results

Here's how the results performed when we were exploring our 
hypotheses

# Discussion

Here's what worked the best, worst, and future areas for research.

# References