### 2018-12-01 report/report_draft.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Identifying Drawings with the Quick Draw Challenge},
            pdfauthor={Blue Team},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Identifying Drawings with the Quick Draw Challenge}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Blue Team}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \date{}
  \predate{}\postdate{}

\begin{document}
\maketitle

We're trying to put together a CNN based on the ``Quick, Draw!'' Kaggle
competition (Google 2018).

\hypertarget{theory-literature-review}{%
\section{Theory \& Literature Review}\label{theory-literature-review}}

CNN with dilation is preferable to LSTM because it's presumably faster.
We can cite some papers here.

\hypertarget{hypotheses}{%
\section{Hypotheses}\label{hypotheses}}

Maybe we should structure our models as experiments. Then talk about
what worked the best. Here's some example hypotheses

\(H_a:\) Prediction will go up when we use a CNN compared to a fully
connected feedforward neural network.

\(H_a:\) Prediction will go up when we use regularization

\hypertarget{data}{%
\section{Data}\label{data}}

We got our data from Google/Kaggle

\hypertarget{results}{%
\section{Results}\label{results}}

Here's how the results performed when we were exploring our hypotheses

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Here's what worked the best, worst, and future areas for research.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-QuickDra10}{}%
Google. 2018. ``Quick, Draw! Doodle Recognition Challenge \textbar{}
Kaggle.'' \url{https://www.kaggle.com/c/quickdraw-doodle-recognition}.

\end{document}