### 2018-12-01 report/report_draft.Rmd
update report

---
title: "Identifying Drawings with the Quick Draw Challenge"
author: "Group 3"
output: 
  pdf_document:
    keep_tex: yes
bibliography: references.bib
---

We're trying to put together a CNN based on the "Quick, Draw!" Kaggle
competition [@QuickDra10].

# Introduction

Our team has based our project idea on the "Quick, Draw!" Kaggle competition 
which classifies user drawings into one of 345 label categories. Many drawings 
were incomplete or failed to match the label. The challenge was to effectively 
build a recognizer that could work with this noisy data. The challenge was unique 
from an algorithmic perspective because the data included both temporal and 
spatial components; areas that have traditionally used two different 
types of Neural Network Architectures. Our team attempted to provide a solution 
using a subset of the data, to the Kaggle problem, while taking a more in depth 
look at potential pitfalls and modeling strategy, without having to use 
extensive computational resources.

# Methods

First, we got our data from Google/Kaggle. CNNs are traditionally known to work
well with spatial data, which prompted us to process the original csv files to 
images of the drawings.

Maybe we should structure our models as experiments. Then talk 
about what worked the best. Here's some example hypotheses

$H_a:$ Prediction will go up when we use a CNN compared to a fully
	connected feedforward neural network.

$H_a:$ Prediction will go up when we use regularization.

$H_a:$ Run neural networks without any activations.

$H_a:$ Try different architectures- number of layers, different (or no) 
activations

$H_a:$ CNN with dilation is preferable to LSTM because it's presumably
faster. We can cite some papers here.

# Results

Here's how the results performed when we were exploring our 
hypotheses

# Discussion

Here's what worked the best, worst, and future areas for research.

# References

# Appendix

Add computer code, plots, and other relevant technical details that will help
me evaluating your work.

# Statement of contribution