### 2018-12-03 proposal/proposal2.Rmd
organize

---
title: "DS 5220 Project Proposal"
author: "Chaitya Shah, Tyler Brown, Karan Gulur Muralidhar, Omair Shafi Ahmed; MSDS 19'"
output: pdf_document
---

**Description of the problem**

Our team is basing our project idea on the "Quick, Draw!" Kaggle competition 
which classifies user drawings into one of 345 label categories. Drawings may 
be incomplete or not match the label. The challenge is to effectively build a 
recognizer that can work with this noisy data. The challenge is unique from an
algorithmic perspective because the data includes both temporal and 
spatial components; areas that have traditionally used two different 
types of Neural Network Architectures. Our team attempts to provide a solution 
using a subset of the data, to the Kaggle problem, while taking a more in depth 
look at potential pitfalls and modeling strategy, without having to use 
extensive computational resources.

**Summary of the data**

Google curated the Quick Draw Dataset as a collection of 50 million drawings
mapped to 345 categories. The drawings were submitted by players of the game
"Quick, Draw!" and captured as timestamped vectors, tagged with metadata such
as the requested drawing category and country code. The format of each drawing
array contains 3-dimensional vectors of strokes, where components are the x,y 
coordinates and time in milliseconds since the first point. We use a preprocessed 
version of the dataset which is simplified and is about tenth the size.
The simplified dataset contains a CSV file for each of the 345 categories.
Google was able to prepocess the data by performing the following steps: (1)
align each drawing to the top-left corner, (2) uniformly scale each drawing, 
(3) resample all strokes with a 1 pixel spacing, and (4) simplify all strokes
using the Ramer-Douglas-Peucker algorithm with an epsilon value of 2.0.

**Methods**

We plan to use Neural Networks for this problem. The problem is unique in
that it has a time series component related to drawing stroke sequences, and
an image recognition component related to the cumulative strokes. Recurrent Neural
Networks (RNN) are generally associated with time series. Convolutional Neural
Networks (CNN) are often associated with image recognition. The inclusion of
time series and pattern recognition components requires us to consider 
tradeoffs related to model accuracy, scalability, and algorithmic complexity.
For example, RNNs have been shown to be less scalable than CNNs due to
computational resource requirements. We have noted that researchers have
attempted to capture benefits of RNNs in CNNs through the use of techniques
such as dilation. The goal of this project is to compare the aforementioned
Neural Network Architectures, and present the findings.

Feature engineering will help us capture the temporal component of our data. We
recognize that choosing appropriate features can reduce the amount of data 
required to train a Neural Network appropriately, as well as reduce the amount
of computational resources required to fit our model.

**Preliminary Results**

Given that Neural Networks can be computationally intensive, one of our
initial assessments was where and how to run these models. We tried 
replicating a CNN which had already worked for a team on Kaggle using the
Google Colab resources. We were unable to replicate their results. We then 
reached out to the Instructor to explore alternatives such as better 
computers or using a subset of our data. We chose to use a subset of the
data to focusing on modeling rather than computational infrastructure.

When visualizing our data, we were able to confirm that the data appears to
be very noisy. We found that "Quick, Draw!" players interpreted categories 
such as "cat" correctly but used different perspectives. One drawing may be 
the face of a cat, another included a side-view with the face, body, and tail.
Our model will need to account for these additional variations.

**References**

Image Based CNN (https://www.kaggle.com/jpmiller/image-based-cnn)