### 2018-12-05 presentation/final.Rmd
started putting together the presentation

---
title: "Quick, Draw! Challenge"
subtitle: DS 5220
author: "Blue team"
output:
  revealjs::revealjs_presentation:
    theme: simple
    highlight: pygments
    center: true
---

## Problem to Solve: Quick, Draw! Image Classification

* We focus on a subset of the Kaggle competition
  - Image classification vs. identifying sequences when drawing

## Our Strategy

* Bias vs. Variance
  - Overfitting 
  - Underfitting
* Multi-Class Classification
  - Accurcy vs. Number of Classes
  - Imbalanced classes
* Neural Network Architecture
  - Fully Connected Neural Network
  - Convolutional Neural Network (CNN)

# Bias vs. Variance

## Seeing the Big Picture

* Neural Networks differ from many other models in
  that bias vs. variance is less of a tradeoff

  - We can fix bias, then separately fix variance

## Overfitting

* overfitting is bad and here's how it happens
* here's how we dealt with it

## Underfitting

* also bad, here's how we deal with it

# Multi-Class Classification

## Why is this necessary?

* We need to classify multiple quickly drawn images

## Accurcy vs. Number of Classes

* Accuracy seemed to go down as the number 
  of classes went up

## Imbalanced classes

* Also bad, used under-sampling

# Neural Network Architecture

## Fully Connected Neural Network

* Designed using Tensorflow without Keras

## Convolutional Neural Network (CNN)

* Designed using Keras with Tensorflow backend

# Summary

## Recommendations

* Neural networks are good for classifying images containing a hotdog

# Thank you! Any questions?