### 2018-12-06 report/report_draft.tex

\documentclass[]{article}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage[margin=1in]{geometry}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Identifying Drawings with the Quick Draw Challenge},
            pdfauthor={Group 3},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\usepackage{longtable,booktabs}
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

%%% Use protect on footnotes to avoid problems with footnotes in titles
\let\rmarkdownfootnote\footnote%
\def\footnote{\protect\rmarkdownfootnote}

%%% Change title format to be more compact
\usepackage{titling}

% Create subtitle command for use in maketitle
\newcommand{\subtitle}[1]{
  \posttitle{
    \begin{center}\large#1\end{center}
    }
}

\setlength{\droptitle}{-2em}
  \title{Identifying Drawings with the Quick Draw Challenge}
  \pretitle{\vspace{\droptitle}\centering\huge}
  \posttitle{\par}
  \author{Group 3}
  \preauthor{\centering\large\emph}
  \postauthor{\par}
  \date{}
  \predate{}\postdate{}

\begin{document}
\maketitle

We're trying to put together a CNN based on the ``Quick, Draw!'' Kaggle
competition (Google 2018).

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Provide a short background of the project (e.g., what kind of question
is to be answered with this work). Provide a short non-technical summary
of your analytical strategy.

Our team has based our project idea on the ``Quick, Draw!'' Kaggle
competition which classifies user drawings into one of 345 label
categories. Many drawings were incomplete or failed to match the label.
The challenge was to effectively build a recognizer that could work with
this noisy data. The challenge was unique from an algorithmic
perspective because the data included both temporal and spatial
components; areas that have traditionally used two different types of
Neural Network Architectures. Our team attempted to provide a solution
using a subset of the data, to the Kaggle problem, while taking a more
in depth look at potential pitfalls and modeling strategy, without
having to use extensive computational resources. With this project, we
explore image classification in depth. We try out various experiments
like fully connected neural network vs CNN, bias vs.~variance tradeoff,
overfitting data purposefully and trying out techniques like
regularization and dropout to fix overfitting and finally, testing
neural network with imbalanced data.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

First, let's talk about data collection. Google initially compressed and
converted the images to .npy files. These .npy files are basically 28*28
images, containing more than 100,000 rows each.

So here are the experiments we performed.

\hypertarget{eda}{%
\subsubsection{EDA}\label{eda}}

We checked the We plotted images from several categories

\(H1_a:\) Prediction will go up when we use a CNN compared to a fully
connected feedforward neural network.

\(H2_a:\) Prediction will go up when we use regularization.

\(H3_a:\) Run neural networks without any activations.

\(H4_a:\) Try different architectures- number of layers, different (or
no) activations

\(H5_a:\) CNN with dilation is preferable to LSTM because it's
presumably faster. We can cite some papers here.

Try out: Overfitting, Bias/variance, imbalance, dropout, pushing
activation

\hypertarget{results}{%
\section{Results}\label{results}}

Here's how the results performed when we were exploring our hypotheses

\begin{longtable}[]{@{}ll@{}}
\toprule
Hypothesis & Result\tabularnewline
\midrule
\endhead
\(H1_a\) & Accept\tabularnewline
\(H2_a\) & Fail to Accept\tabularnewline
\(H3_a\) & Fail to Accept\tabularnewline
\(H4_a\) & Accept\tabularnewline
\(H5_a\) & Fail to Accept\tabularnewline
\bottomrule
\end{longtable}

\hypertarget{fully-connected-neural-network}{%
\subsection{Fully Connected Neural
Network}\label{fully-connected-neural-network}}

Here's the run down on it and how it relates to the hypotheses.

\hypertarget{convolutional-neural-network}{%
\subsection{Convolutional Neural
Network}\label{convolutional-neural-network}}

same thing for the CNN. TOOD: can we fold in some of the specializations
into this section??

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Here's what worked the best, worst, and future areas for research.

The model built using CNN was better than human accuracy, which was in
turn better than the model built using fully connected network.

\hypertarget{references}{%
\section{References}\label{references}}

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

Add computer code, plots, and other relevant technical details that will
help me evaluating your work.

\hypertarget{statement-of-contribution}{%
\section{Statement of contribution}\label{statement-of-contribution}}

Everyone worked on everything.

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-QuickDra10}{}%
Google. 2018. ``Quick, Draw! Doodle Recognition Challenge \textbar{}
Kaggle.'' \url{https://www.kaggle.com/c/quickdraw-doodle-recognition}.

\end{document}