### 2018-12-07 report/draft.tex
moved report over to latex with minimal reformatting

%
% PPUA 5262 Final
%
\documentclass[12pt]{article}

%
% Packages
%
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{clrscode3e}
\usepackage{enumerate}
\usepackage{graphicx}
\usepackage[table]{xcolor}% http://ctan.org/pkg/xcolor
\usepackage{caption}
\usepackage{subcaption}

%
% Package settings
%

%% begin.rcode setup, include=FALSE
% library(knitr)
%
% opts_chunk$set(fig.path='figure/latex-', cache.path='cache/latex-')
%% end.rcode

%
% Document Settings
%
\setlength{\parskip}{1pc}
\setlength{\parindent}{0pt}
\setlength{\topmargin}{-3pc}
\setlength{\textheight}{9.5in}
\setlength{\oddsidemargin}{0pc}
\setlength{\evensidemargin}{0pc}
\setlength{\textwidth}{6.5in}

\title{Quick, Draw! Image Classification}
\author{Group 3}

%
% Image directory location
%
\graphicspath{ {./imgs/} }

\begin{document}

\maketitle

\section{Introduction}

Many recent industry innovations have included neural networks. An
important research area within neural networks has been image classification.
One of the best techniques to classify images is Convolutional Neural
Networks (CNN). We have sought to better understand industry best practices
by exploring different image classification techniques with neural networks.
Our exploration starts by manually performing image classification and
noting the level of human accuracy. We then use a fully connected neural
network to perform image classification before implementing a CNN. We
implemented options for tuning and debugging the neural network while
implementing each method during our process. We were able to develop an
intuition for image classification best practices using neural networks
after our research.

Many supervised machine learning methods often differ from those employed 
when constructing neural networks. These supervised machine learning
methods, such as logistic regression or decision tree classification,
require feature engineering to successfully classify an image. Models
requiring feature engineering rely on considerable effort by subject matter
experts to achieve reasonable accuracy. Neural networks allow for an
approach which does not require feature engineering to achieve similar or
better accuracy. Industry engineering efforts regarding feature engineering
can be shifted to neural networks  in some cases such as image
classification. Moreover, the availability of cheap computing resources
sped up a shift towards the use of neural networks.

When considering modeling approaches, we examine both fully connected and 
convolutional neural networks. A fully connected neural network is known to 
be inefficient at classifying large images. These inefficiencies in a fully 
connected neural network partly arise because every neuron in the $\ell-1$
layer must be connected to the $\ell$th layer. This requires a series
of matrix multiplications and additions across each layer $\ell$.
Alternatively, a CNN allows for more efficient computations by: (1) efficient
and automatic feature generation using local connectivity and parameter
sharing of convolution operations, (2) dimensionality reduction using pooling
layers. In this project, we attempt to understand these topics in depth
from an empirical stanpoint in relation to image classification using
Google's ``Quick, Draw!'' dataset.

\section{Methods}

We generate a series of hypotheses to inform our analysis.

\subsection{Hypotheses}

We start off with hypothesis that we learnt from theoretical literature. We
then perform several experiments to test the validity of these hypothesis and
report out our findings. Some of our hypothesis is as follows:

1) Fully connected neural network will struggle to get reasonable accuracy.
In contrast CNN’s should perform better and faster than a Fully connected
neural network.

2) Training a deep neural network with limited data will result in
overfitting. Conversely, adding more data to a deep neural network is an
effective way to manage variance.

3) Imbalance class prediction can be improved using data augmentation.

4) Increasing the number of categories will cause a significant drop in the
performance of the neural network. Increasing the overall data points for the
network to be trained on should help.

\subsection{Data}

Google originally collected the data from users by explicitly asking them to
draw objects like forks, hotdogs, etc. The original dataset is 70 GB in size
but Google also provides a simplified version of the dataset- consisting only
of the final images. We used these '.npy' files to train all of our models.
These numpy bitmap files consist of more than 100,000 rows each and 784
features, since they are basically $28 \times 28$ images. Samples of these
images are shown in Figure \ref{fig:quickImages}. The subset of this
dataset sampled varies from experiment to experiment. More details can be
found in the experiments mentioned below.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.5]{fig1}
  \end{center}
  \caption{Sample images from Quick, Draw! dataset}
  \label{fig:quickImages}
  \end{figure}

\section{Results}

\subsection{Hypothesis 1}

Hypothesis 1: Fully connected neural network will struggle to get reasonable
accuracy. In contrast CNN’s should perform better and faster than a Fully
connected neural network.

In order to perform this experiment we built a fully connected neural network
using tensorflow. Our dataset sample selected for this experiment had 10
categories: fish, fork, hotdog, flamingo, airplane, alarm clock, baseball bat,
bicycle, dolphin, elephant. We had 100,000 examples in total, with 10,000
example for each of the aforementioned categories. Our train:test split was
chosen to be 80:20, an industry standard.

Our initial objective was to build a model that overfit our given dataset,
with the idea that once our model is complex enough to overfit the training
data, we could regularize it in order to achieve a more optimal fit.

After experimenting with various architectures, we narrowed down to a 3 layer
model as specified in Figure \ref{fig:fcnn}.

\begin{figure}[h]
  \begin{center}
    \includegraphics[scale=0.75]{fig2}
    \end{center}
  \caption{Fully Connected Neural Network Architecture}
  \label{fig:fcnn}
  \end{figure}

The training set accuracy observed was 93\%, while the test accuracy was at
73\%, which clearly indicated an overfit in the model, as designed.

The model took 1500 epochs to converge taking around ~45 minutes on a machine with 13 GB of RAM, Intel Xeon CPU 2.3 GHz (1 core, 2 threads), 1xTesla K80 GPU having 2496 CUDA cores with 12 GB GDDR5 VRAM. In order to manage variance, we attempted regularizing using dropouts. Unfortunately, that effort was futile. We then doubled the amount of data sampled to train the model, eventually helping us achieve training set accuracy of 86\% and test set accuracy of 82\%. Our
learning rate was specified as 0.001.

While our initial hypothesis assumed that a regular fully connected network would struggle to achieve a reasonable accuracy, our experiment proved our hypothesis as invalid by achieving a commendable accuracy. Having said that, the model did take a longer to converge than expected.

As for the CNN, we started off with the same approach with attempting to overfit the model using convolution operations and a fully connected layer. Since our objective is to categorise 10 categories, softmax was chosen as the activation function for the final layer, while every other neuron had ReLU as the activation function owing to ReLU’s faster convergence.

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.5]{fig3}
    \end{center}
  \caption{CNN Architecture}
  \label{fig:cnnArch}
\end{figure}

Loss Function: Categorical Cross Entropy [TO DO: Why?]
Optimizer: [TODO: Why?]
Number of Trainable Parameters: 320,890

The number of clearly being much greater than the number of observations used to train the model, we need to reduce the number of parameters vis-a-vis number of observations. From the available strategies to do so, we incorporated MaxPooling [TO DO: why?] post each convolution operation, in order to extract only the most significant features. As as a consequence, this reduced the volume of the output, giving the model a smaller memory footprint as well as a smaller number of trainable parameters.

\begin{figure}
  \begin{center}
    \includegraphics[scale=0.5]{fig4}
  \end{center}
  \caption{CNN Architecture with Reduced Parameters}
  \label{fig:cnnArchPool}
\end{figure}

The observable training accuracy is 0.90, whereas the validation accuracy is 0.89. The model converged after 10 epochs, which took ~ 3 minutes.

Some of the other unsuccessful strategies used to decrease the number of parameters and to regularize the network were strides and dropouts which sacrificed way too much in the way of accuracy and thus had to be dropped. Since most of our drawings had most of their information in the center of the image, padding, in theory wouldn’t be of much help. However, actually implementing padding improved our accuracy by 1\% at a cost of ~6000 parameters.

Conclusion

A slight (~5\%) improvement in the accuracy of a CNN when compared to a fully connected network, keeping the number of parameters constant, was observed. However, the CNN happened to be around 13 times faster in practice. This delta in accuracy and speed is likely to be amplified as we scale the model up with respect to model complexity or data.

\subsection{Hypothesis 2}

Hypothesis 2: Training a deep neural network with limited data will result in overfitting. Conversely, adding more data to a deep neural network is an effective way to manage variance.

With us reaching an optimal model for the dataset, we decided to study the effect of reduction in the sampled dataset and its relationship with overfitting.

\subsection{Hypothesis 3}

We tried to show the importance of having equal partitions of data to train a machine learning model. Imbalanced classes can make a model biased toward the majority class.

\section{Discussion}

\bibliographystyle{unsrt}
\bibliography{references}

\end{document}